---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
BMisc::source_all("~/Dropbox/pte/R/")
BMisc::source_all("~/Dropbox/ppe/R/")
load("~/Dropbox/ppe/data/covid_data.rda")
library(ggplot2)
```
# Panel Treatment Effects (pte) Package

The `pte` package compartmentalizes the steps needed to implement estimators of group-time average treatment effects (and their aggregations) in order to make it easier to apply the same sorts of arguments outside of their "birthplace" in the literature on difference-in-differences.

This code is lightweight, only works for balanced panels, and has minimal error checking.  That said, it should be useful projects that build on top of group-time average treatment effects in order to deliver estimates of causal effects in panel data settings.

Here are a couple of examples:

## Example 1: Difference in differences

*Note:* This is not the recommended way of doing this (as there is very little error handling, etc. here), but it is rather a proof of concept.
```{r}
library(did)
data(mpdta)
did_res <- pte(yname="lemp",
           gname="first.treat",
           tname="year",
           idname="countyreal",
           data=mpdta,
           subset_fun=two_by_two_subset,
           attgt_fun=did_attgt,
           xformla=~lpop) 

summary(did_res)
ggpte(did_res)
```
What's most interesting here, is that the only "new" code that needs to be writte is in the `did_attgt` function [available here](https://github.com/bcallaway11/pte/blob/master/R/attgt_functions.R).  You will see that this is a very small amount of code.

## Example 2: Policy Evaluation during a Pandemic

This comes from [Callaway and Li (2021)](https://arxiv.org/abs/2105.06927).

```{r}
# formula for covariates
xformla <- ~ current + I(current^2) + region + totalTestResults
```

```{r echo=FALSE, results=FALSE, warning=FALSE}
# drop some data as in paper
trim_id_list <- lapply(c(10,15,20,25,30),
                       did::trimmer,
                       tname="time.period",
                       idname="state_id",
                       gname="group",
                       xformla=xformla,
                       data=covid_data,
                       control_group="nevertreated",
                       threshold=0.95)
time_id_list <- unlist(trim_id_list)

# unique(subset(covid_data, state_id %in% time_id_list)$state)
covid_data2 <- subset(covid_data, !(state_id %in% time_id_list))
```


```{r}
covid_res <- pte(yname="positive",
                 gname="group",
                 tname="time.period",
                 idname="state_id",
                 data=covid_data2,
                 subset_fun=two_by_two_subset,
                 attgt_fun=covid_attgt,
                 xformla=xformla,
                 max_e=21,
                 min_e=-10) 

summary(covid_res)
ggpte(covid_res) + ylim(c(-1000,1000))
```

What's most interesting here is just how little code needs to be written here.  The only new code required is the `ppe::covid_attgt` function which is [available here](https://github.com/bcallaway11/ppe/blob/master/R/covid_attgt.R), and, as you can see, this is very simple.

